The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-06-28 brains-conclusion.tex
feat/related-work: major rework of the "related work" part

>>>>

Attraction can also be defined on any number of dimensions. It can be on the structure, on anything, so saying there could be an attraction while not specifying the dimension is really meaningless. What's more important is to look at a specific dimension, and see if there's attraction on that one, as we did here for features.

We could've done also on semantic grouping, predicting the new word based on semantic similarity (or on frequent dyads, i.e. collocation with previous word the same way Zaromb et al. 2006 explain PLIs), and predicting the disappearing word based on the cost of doing such a substitution (lower cost -> higher prob of substitution). The point is, there's decades and many fields of psycholinguistics, and we can connect each of them with this question.

All of this is possible with our software that we published.

ADDTHIS:
Taking context into account is more than what we did. For instance, building on Zaromb 2006, you could imagine that the substitutions appear because the word preceding the substituted one appears in another dyad a lot more than this one, triggering a substitution (Zaromb's associative vs. contextual retrieval processes in recall).

ADDTHIS:
Soooooo... following literature on word lists (Zaromb and DRM):
- we could predict the new word in substitution? Take the strongest average association to words in the sentence.
- we could predict substituted word? Take the word following the one that triggers the new word.
Problems:
- we probably won't find the exact word, but one similar to it (even Zaromb can't predict the exact word, they don't try, they just check it comes from the right list). How to evaluate that?
- if computing LSA/LDA on the corpus (which probably isn't adapted because of the short sentence nature of the data -> topics = quote families), it's tautological unless you suppose substitutions have a negligible effect.
Explain that to the reviewer.

Again justify our approach w/ features by the shape of our data: few substitutions per cluster (avg. 9), and substitutions on relatively few clusters overall -> opposite situation to a few lists repeated through 50 subjects where frequency of transition has meaning. Here for each word, it's nearly one shot. So we have to categorize words (by using low number of features -> known features best) to get frequencies. From there, you can predict many many things, so better describe.

### 2016-06-28 brains-introduction.tex
feat/related-work: major rework of the "related work" part

\begin{new}
Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
%the reciprocal influences of cognition and culture in meaningful ways.
While it has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g. Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.

@@

The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how the theory of memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
\Citet{mesoudi_multiple_2008} have discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
\Citeauthor{kirby_cumulative_2008} \citeyearpar{kirby_cumulative_2008,cornish_systems_2013} have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
Such transmission chain experiments have also been explored in non-human primates by \citet{claidiere_cultural_2014}.

@@

The theory makes an additional strong hypothesis, which this paper focuses on:
as transformations accumulate, some representations evolve to be very stable and spread throughout an entire society without changing any more (they are called "cultural representations", because they characterize a given culture).
This process should manifest itself as attractors (called "cultural attractors") in the dynamical system that models cultural evolution, that is:
there should be areas of the representation space where any cognitive effect in transformation brings representations closer to a given stable asymptotic point.\footnote{
Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
}

This hypothesis, a cornerstone of the theory because of the intelligibility it gives to cultural evolution, has been hard to test in concrete situations:
quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
One approach, as mentioned above, has been the meta-analysis of large bodies of anthropological studies (see \citealp{miton_universal_2015}, for instance).
This paper exemplifies a second approach, taking advantage of the ever-increasing avalanche of available digital footprints since the 2000's.
Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly.
For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in the email network~\citep{liben-nowell_tracing_2008};
\citet{cointet_socio-semantic_2009} have described the reciprocal influence between the social network topology and the distribution of issues;
\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} have studied the characteristics of particularly memorable quotes that circulate in those networks.
We believe those works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.

@@

\todo{
\#15: add
(1) sentence recall (Potter \& Lombardi),
(2) false memories (Deese),
(3) subjective organization (Tulving, Zaromb),
(4) working memory and attention (Jefferies),
(5) iterated learning (Kirby)
}

@@

%The relevant literature on \emph{public representation dynamics} features two main streams.
%On one hand, we find studies of the macroscopic \emph{social diffusion} of public representations, describing for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.
%These studies are relatively independent from anthropology and cognition and are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis).
%Without necessarily relying on specific social science theories, this research stream is of interest for its use of large textual corpora in studying cultural dynamics.

The practical study of the transformation of public representations has emerged only recently.
For one, models involving evolution and representations to study the notion of ``cultural attractor'' have appeared only a few years ago (see \citealp{Claidiere07}, and \citealp{claidiere2014darwinian}, as well as a hybrid empirical-theoretical protocol in \citealp{maccallum2012evolution}).
Among the empirical approaches, some of the most relevant studies to date consist in a series of papers investigating \emph{quotation} transformations in a large corpus of US blog posts, initially collected and studied by \citet{Leskovec09} and further analyzed by \citet{Simmons11} and \citet{omod-mult}.
%They exhibit several types of regularities and propose diffusion-transformation models of the evolution of quotations, which may nonetheless appear to be relatively simplistic from a cognitive viewpoint. One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations. However, these studies address transformations by focusing on the properties of the source of the quotation (\hbox{e.g.} news outlet {vs.} blog), or the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), rather than the very cognitive-level features which may determine or, at least, influence these transformations.
One of the main observations in these works is that even for quotations, a type of public representation that should be among the most stable, it is still possible to witness significant transformations. They essentially examine the effect of some properties of the quotation source (\hbox{e.g.} news outlet {vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus). Some diffusion-transformation models have been proposed, yet the very cognitive features which may determine or, at least, influence these transformations, are overlooked; which may appear to be relatively unsatisfying from a cognitive viewpoint.

At this level, we have to turn to the broader psycholinguistic literature which provides one of the main cognitive foundations for public representation evolution by studying the influence of word features on the ease of recall.
This field is well developed and details the impact that classical psycholinguistic variables such as word frequency~\citep[see][for a review]{Yonelinas02}, age-of-acquisition~\citep{Zevin02}, number of phonemes or number of syllables~\citep[see for instance][]{Rey98,nick-diss}, have on this type of task.

Less classical linguistic variables, based on the study of semantic network properties, have recently started to be used, in the context of connectionism and its normative processual models~\citep[see for instance][]{collins1975spreading}.
Let us mention four interesting studies on that matter, which demonstrate in a strictly \emph{in vitro} framework and at the vocabulary level that properties computed on a word network are important factors for the cognitive processes and reproduction of those words.
First, \citet{Griffiths07} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the %Latin
alphabet. The authors show that there exists a link between the ease of recall of words and one of their semantic features, namely their authority position (pagerank) in a language-wide semantic network built from external word association data.
\citet{austerweil2012human} further develop this idea by showing that random walk on such a semantic network, that is the exact process measured by the pagerank index, gives a parsimonious account of some semantic retrieval effects (namely, related items being retrieved together).
A third psycholinguistic study by \citet{Chan10} shows, in a picture-naming task, that words are produced faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network (which, again, is  defined from external phonological data).
\citet{nelson2013activation}, finally, show the importance of clustering coefficient in a semantic network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free assocation).

On the whole, the current psycholinguistic state-of-the-art seems to hint towards two antagonistic types of results.
On one hand, part of the literature tends to show that recall is easier for the least ``awkward'' words; those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
On the other hand, when the task consists in recognizing a specific item in a list, ``awkward'' words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{Yonelinas02}.
The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
We also aim here at shedding some light on this debate, considering oddness as a dimension of the purported fitness of utterances.

>>>>

\begin{new}

TOO GENERAL OR GIVES THE IMPRESSION OF TOO HIGH AMBITIONS. COMPARE TO OTHER INTROS TO JUST SAY WE'LL TALK ABOUT THIS.

Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
%the reciprocal influences of cognition and culture in meaningful ways.
While this has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g. Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.

@@

The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how the theory of memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
\Citet{mesoudi_multiple_2008} have discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
\citeauthor{kirby_cumulative_2008}~(\citeyear{kirby_cumulative_2008,cornish_systems_2013}) have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
Such transmission chain experiments have also been explored in non-human primates by \citet{claidiere_cultural_2014}.

@@

The theory makes an additional strong hypothesis, which this paper focuses on:
as transformations accumulate, some representations evolve to be very stable and spread throughout an entire society without changing any more (they are called "cultural representations", because they characterize a given culture).
This process should manifest itself as attractors (called "cultural attractors") in the dynamical system that models cultural evolution, that is:
there should be areas of the representation space where cognitive effects in transformations bring representation closer to a given stable asymptotic point.\footnote{
Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
}

This hypothesis, a cornerstone of the theory because of the intelligibility it gives to cultural evolution, has been hard to test in concrete situations as quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
One approach, as mentioned above, has been the meta-analysis of large bodies of anthropological studies (see \citealp{miton_universal_2015}, for instance).
This paper exemplifies a second approach, taking advantage of the ever-increasing avalanche of available digital footprints since the 2000's.
Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly.
For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in the email network~\citep{liben-nowell_tracing_2008};
\citet{cointet_socio-semantic_2009} have described the reciprocal influence between the social network topology and the distribution of issues;
\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} have studied the characteristics of particularly memorable quotes that circulate in those networks.
We believe those works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.

@@

@@

The study of \newtext{cultural evolution on the part of cognitive science} emerged only recently.
\newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
Critics like \citet{ingold_trouble_2007}, however, have noted that such quantitative use of works from the social sciences approach risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
}
A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
Transmission chains, in particular, have been used extensively to study the evolution of human language~\citep[see][for a review]{tamariz_cultural_2016}.
Other recent examples include studies of the evolution of simple audio loops through consumer preference~\citep{maccallum_evolution_2012}, the emergence of structure in visual patterns transmitted by baboons~\citep{claidiere_cultural_2014}, and the amplification of risk perception through chains of casual conversation~\citep{moussaid_amplification_2015}.}

\newtext{Research on online content points to a third approach to this question.}
By investigating the transformations of quotations in a large corpus of US blog posts and online news stories initially collected and studied by \citet{leskovec_meme-tracking_2009}, \citet{simmons_memes_2011} and later \citet{omodei_multi-level_2012} show that even for quotations, a type of public representation that should change the least when transmitted on the Web, it is still possible to witness significant transformations.
These studies focus on the influence of the quotation source (e.g. news outlet vs. blog) or of the surrounding public space (e.g. quotation frequency in the corpus), and suggest diffusion-transformation models to capture the dynamics of the population of quotations.
But the cognitive features which may determine or, at least, influence these transformations, are overlooked.
\newtext{Yet cognitive and linguistic features have been used in studies not involving transformation: \citet{danescu-niculescu-mizil_you_2012}, for instance, show that particularly memorable quotations (taken from movie scripts in this case) use more distinctive words and have more common syntax than less memorable quotations; they are also the quotes that adapt best to new contexts of use.
One source of ideas to study the transformations of such quotes, then, might be the psycholinguistic literature studying word and sentence recall.}

\begin{new}

\citet{potter_regeneration_1990} suggest that immediate recall of sentences is based on the retention of an unordered list of words which is then regenerated as a sentence at the moment of production.
Priming recall with other words can lead to replacement in the recalled sentence if the primed words support the overall meaning of the sentence.
Regenerated syntax can also be influenced by priming recall with another syntactic structure~\citep{potter_syntactic_1998}, or with verbs whose category constraints call for a different structure~\citep{lombardi_regeneration_1992}.

Recall of word lists provides a situation that is easier to explore fully compared to recall of complete sentences, and has been extensively studied.
In particular, the Deese, Roediger, and McDermott paradigm (introduced by \citealp{deese_prediction_1959}, and later popularized by \citealp{roediger_creating_1995}) has shown that it is possible to construct lists of words which reliably create the false memory of an external word related to those in the list.
This is done by using lists of words produced by free association from the target intrusion word;
the intruding recall then happens with probability nearly proportional to the average strength of semantic association between the intruding word and the words in the list.
A sizable literature studies this type of task with varying complexities in the design of the lists, a good review of which is given by \citet{zaromb_temporal_2006}.
One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which subjects recall a list~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}
The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006};
indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations (contributing to associative information retrieval) of subjects is at the center of many of these studies.

These effects don't transpose simply to sentence recall however, as not only syntax but also effects of attention come into play for both retrieval and encoding.
\citet{jefferies_automatic_2004}, for instance, show that attention is central to the encoding and retention of unrelated propositions, on top of more automatic syntactic and semantic processes.
This involvement of executive resources also seems to contribute to the much greater memory span subjects exhibit for sentences compared to word lists~\citep[see][again, for more details]{jefferies_automatic_2004}.

Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited when the data consists of a high number of measures over a smaller number of lists (in which case it makes sense to ask e.g. what proportion of intrusions come from prior lists).
As is explained further down however, our data set is shaped the opposite way: a great number of sentences, with only very few to no measures at all on each sentence.
}
If a cognitive bias in the substitution of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.

Lexical features, then, are obvious well-studied word measures that can be analyzed in aggregate.
Indeed word frequency~\citep[see][for a review]{yonelinas_nature_2002}, age-of-acquisition~\citep{zevin_age_2002}, number of phonemes~\citep[see for instance][]{rey_phoneme_1998,nickels_dissociating_2004}, and phonological neighborhood density~\citep{garlock_age--acquisition_2001} to name a few, all have known effects on word recognition or production.
More complex features based on word networks built from free association or phonological data have also been analyzed:
\citet{nelson_how_2013} for instance, show the importance of clustering coefficient in such a network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free association).
\citet{chan_network_2010} show that pictures are named faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network.
\citet{griffiths_google_2007} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the alphabet.
The authors show that there is a link between the ease of recall of words and their authority position (pagerank) in a language-wide semantic network built from external word association data~(\citealp{austerweil_human_2012}, further develop this tool to give a parsimonious account of the fact that related words are often retrieved together from memory).

On the whole, research on lexical features hints towards two antagonistic types of effects.
On one hand, part of the literature shows that recall is easier for the least "awkward" words;
those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
On the other hand, when the task consists in recognizing a specific item in a list, "awkward" words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{yonelinas_nature_2002}.
The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
We also aim to shed some light on this debate, considering oddness as a dimension of the purported fitness of utterances.
